TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-07-30 paper/draft.tex
wording

M dwarfs have enourmous potential for understanding structure and
formation on both Galactic and exoplanetary scales
with their properties and compositions.
However, insufficient physical assumptions behind atmospheric models
have limited abilities at reproducing spectral features in stars of the 
coolest temperature regimes ($T_{\rm eff} < 4200\,$K) and fully
exploiting the information content and statistical extent of
current and upcoming large-scale spectroscopic surveys.
Here we present a catalog of spectroscopic temperatures, metallicities,
spectral types for 5,875 M dwarfs in the \apogee\ survey using
\thecannon: a flexible, data-driven spectral-modeling and
parameter-inference framework demonstrated to estimate
stellar-parameter labels ($\teff$, $\logg$, $\feh$, and detailed
abundances) to high precision.
Using a training sample of 87 M dwarfs with optically derived labels spanning 
$2860 < \teff < 4130$\,K callibrated with bolometric temperatures, 
$-0.5 < \feh < 0.5$\,dex callibrated with FGK wide binary metallicities, 
we train a two-parameter model with predictive accuracy (in cross-validation) 
to 77\,K and 0.09\,dex respectively.
We also train a one-dimensional spectral type model using 51 M dwarfs
with known SDSS optical spectral types ranging M0 to M9, to predictive accuracy of
0.9 types.
\color{gcolor}{From the overlap of \textsl{APOGEE DR14} and \textsl{Gaia DR2} we
apply our models and compare our \cannon\ derived $\teff$ against color--temperature
relations and \cannon\ derived $\feh$ against isochrone models in
\gaia\ color-magnitude space.
Finally we use our results to
estimate stellar radii from absolute magnitudes and \cannon\ temperatures [AND WE FIND WHAT??].}\color{black}
\end{abstract}

@@

>>>>

M dwarfs have enourmous potential for understanding structure and
formation on both Galactic and exoplanetary scales
with their properties and compositions.
However, current atmospheric models
have limited abilities at reproducing spectral features in stars of the 
coolest temperature regimes ($T_{\rm eff} < 4200\,$K) and fully
exploiting the information content and statistical extent of
current and upcoming large-scale spectroscopic surveys.
Here we present a catalog of spectroscopic temperatures, metallicities,
spectral types for 5,875 M dwarfs in the \apogee\ survey using
\thecannon: a flexible, data-driven spectral-modeling and
parameter-inference framework demonstrated to estimate
stellar-parameter labels ($\teff$, $\logg$, $\feh$, and detailed
abundances) to high precision.
Using a training sample of 87 M dwarfs with optically derived labels spanning 
$2860 < \teff < 4130$\,K callibrated with bolometric temperatures, 
$-0.5 < \feh < 0.5$\,dex callibrated with FGK wide binary metallicities, 
we train a two-parameter model with predictive accuracy (in cross-validation) 
to 77\,K and 0.09\,dex respectively.
We also train a one-dimensional spectral type model using 51 M dwarfs
with known SDSS optical spectral types ranging M0 to M9, to predictive accuracy of
0.9 types.
\color{gcolor}{From the overlap of \textsl{APOGEE DR14} and \textsl{Gaia DR2} we
apply our models and compare our \cannon\ derived $\teff$ against color--temperature
relations and \cannon\ derived $\feh$ against isochrone models in
\gaia\ color-magnitude space.
Finally we use our results to
estimate stellar radii from absolute magnitudes and \cannon\ temperatures [AND WE FIND WHAT??].}\color{black}
\end{abstract}

@@

As expected, the Toomre diagram in Figure \ref{fig:toomre} shows that higher metallicity sources in the sample are concentrated in low velocity space corresponding to the thin disk population, while the thick disk population contains a higher concentration of lower metallicity sources.
